---
title: "Covid-19 Vaccine Effects Case Control EAVE - Severe Covid following a positive Covid test"
author: "Chris Robertson, Emily Moore, Jim McMenamin, Aziz Sheikh"
date: ' `r format(Sys.Date(), "%d %B %Y")` '
output: html_document
---

```{r setup, include=FALSE,echo = FALSE, warning=FALSE, error=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, error=FALSE)

# 01 Setup ####
#Libraries
library(plyr)
library(tidyverse)
library(survival)
#library(htmltools)
#Load data

Location <- "/conf/"  # Server
#Location <- "//isdsf00d03/"  # Desktop
project_path <- paste0(Location,"EAVE/GPanalysis/progs/CR/Vaccine")

results <- readRDS(paste0(project_path,"/output/temp/CR_cc_summary_icu_death.RDS")) 
results$Vacc.Status <- factor(results$vacc_status, levels = c("uv","v1_0:6","v1_7:13",  "v1_14:20", "v1_21:27","v1_28:34","v1_35:41", "v1_42+", "v2_0:6","v2_7+"))

upper_plot_limit <- 5
lower_plot_limit <- 0.05

unique_groups <- unique(results$id) #this is an integer id number
n_groups <- length(unique_groups)
```

## Introduction

These estimates are based upon a nested case control design where vaccination records are linked to Ecoss, NRS Death Records, SMR01 and Rapid for ICU Admission for COVID 19 or Death following a positive test, within 28 days.

Cases are all individuals with severe covid - ICU admission of covid death (either on death certificate or any death within 28 days of a positive test).  10 controls were selected at random, matched on age and gender, from all individuals who had not had severe covid up to the date of the severe outcome in the case. 

The Ecoss data includes all PCR test results from both pillar 1 (NHS labs) and pillar 2 (Lighthouse government labs).

Data were extracted for all tests between 8th December 2020 and `r format(Sys.Date()-2, "%d %b %Y")`.  

The linkage to the national vaccination record within GP practices was achieved using the community health index (EAVE Study link number).  

Only vaccinations up to and including the last date of an event were included.  Everyone vaccinated after the date of the last event among the cases was treated as unvaccinated in this analysis.

Individuals who have previously tested PCR positive for covid 19 prior to 8th December 2020 are excluded from this analysis

The EAVE-II study design was used for clinical risk group information fro GP records.  Also


## Statistical Methods

A matched case control design was used to estimate the odds of the event. 

This analysis includes the Pfizer/BioNTech vaccine or both vaccines combined. When considering type specific vaccine effects for the Pfizer/BioNTech vaccine, individuals who had received the AZ vaccine were excluded.

The following time periods between vaccination and date of event are considered: post dose 1: 0--6 days, 7-13 days, 1 week : 14-20 days,  2w : 21-27 days, 3w : 21-28 days, up to and 6+ weeks; and post dose 2: 0-6 days , 7+ days.  

Analysis was by conditional logistic regression.  Further adjustment was made for and deprivation and risk groups.  Calendar time, in weeks, was included as a stratification variable as the background risk of hositalisation varied over time.  

In the plots the lower and upper limits of the confidence intervals are truncated at `r lower_plot_limit` and `r upper_plot_limit` respectively.




```{r table_1, , results='asis', include=TRUE}
options(knitr.kable.NA = '')

for (i in unique_groups) { 
  
  #i <- unique_groups[1]
  
  results_i <- filter(results, id==i)
  
  cat("\n##",as.character(unique(results_i$title)), " \n") 

z_res <- results_i %>% dplyr::select(Vacc.Status, N, R , HR_raw:UCL_raw, HR_adj:UCL_adj) 

z_res_1 <- z_res %>% 
    mutate( UCL_raw = ifelse(UCL_raw > 100*upper_plot_limit , NA, UCL_raw),
            UCL_adj = ifelse(UCL_adj > 100*upper_plot_limit , NA, UCL_adj)
             )

 z_tab <- kableExtra::kbl(z_res_1, 
                           caption = "Numbers and Hazard Ratios", digits=c(NA,0,0,rep(2,6)), 
                           col.names = c("Vaccine Group","N","Number Events",
                                         "HR","LCl","UCL","HR","LCl","UCL" )) %>%
    kableExtra::kable_classic(full_width = F, html_font = "Cambria", position = "left") %>%
    kableExtra::column_spec(2:3, width="6em") %>% 
    kableExtra::column_spec(c(4,7), width="5em") %>% 
    kableExtra::add_header_above(c(" " = 3, "Unadjusted" = 3, "Adjusted" = 3))



print(z_tab)

z_res <- z_res %>% 
    mutate( UCL_raw = ifelse(UCL_raw > upper_plot_limit , upper_plot_limit, UCL_raw),
            UCL_adj = ifelse(UCL_adj > upper_plot_limit , upper_plot_limit, UCL_adj),
            HR_raw = ifelse(HR_raw > upper_plot_limit , upper_plot_limit, HR_raw),
            HR_adj = ifelse(HR_adj > upper_plot_limit , upper_plot_limit, HR_adj), 
            LCL_raw = ifelse(LCL_raw < lower_plot_limit, lower_plot_limit, LCL_raw),
            LCL_adj = ifelse(LCL_adj < lower_plot_limit, lower_plot_limit, LCL_adj),
            HR_raw = ifelse(HR_raw < lower_plot_limit, lower_plot_limit, HR_raw),
            HR_adj = ifelse(HR_adj < lower_plot_limit, lower_plot_limit, HR_adj) )
  


g1 <- z_res %>% ggplot(aes(Vacc.Status, y= HR_adj)) + geom_point() +
  geom_errorbar(aes(ymin = LCL_adj, ymax = UCL_adj), width = 0.2) + geom_hline(yintercept=1)+
  labs(x="days from vaccination dose", y= "Adjusted Hazard Ratio", 
       title = paste(unique(results_i$title))) + scale_y_log10() + 
  theme_classic() 

print(g1)

  cat("  \n")
  cat("  \n")
  cat("  \n")
  
}

```


##EAVE-II

We linked individual patient-level data from all primary, secondary, mortality and virological/serological testing data in Scotland. We used this national dataset to investigate the temporal progression of COVID-19 in the Scottish population and the development of COVID-19 morbidity and mortality in individuals. 

###Study design

Almost all the residents in Scotland (including children) are registered with primary care, which provides a comprehensive array of health-care services (free at the point of care), including the issuing of prescriptions for medications. Access to secondary care is typically through a general practitioner (GP) based within a primary care practice or via the emergency department (ED) or out-of-hours primary care services. During the acute phase of the pandemic, community based COVID-19 hubs (a GP-led service designed to segment patients and reduce the risk of nosocomial infections) were established. We used anonymised individual patient-level data from general practices, hospitals, death registry, and reverse-transcriptase polymerase chain reaction (RT-PCR) testing (including NHS and UK Government testing centre data) to investigate the changing dynamics of COVID-19 during 2020.

We developed an open cohort drawn from patients registered with any primary care practice in Scotland from March 1st 2020. 

###Datasets
We used data from all 929 Scottish primary care practices. Clinical data collected by primary care in Scotland have consistently been shown to be of high quality (90% completeness and accuracy) and their value for epidemiologic research has been repeatedly demonstrated. These were linked to other datasets including the Electronic Communication of Surveillance in Scotland ECOSS (all virology testing national database), the Scottish Morbidity Record (hospitalisation including ICU stay), and National Records Scotland (death certification).


### Funding

EAVE II is funded by the Medical Research Council and the National Institute for Health Research Health Technology Assessment Programme, with the support of BREATHE – The Health Data Research Hub for Respiratory Health, funded through the UK Research and Innovation Industrial Strategy Challenge Fund and delivered through Health Data Research UK. Additional support through the Scottish Government Director General Health and Social Care


